---
title: "Presentation"
author: "Rose Determan"
date: "10/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("wrangling_code.R")
source("visualization_function.R")
```

I chose to examine **literacy rates and poverty rates** from the
[gapminder dataset](https://www.gapminder.org/data/).    

**Literacy Rates:**  
Gapminder defines literacy rate as the percentage of people ages 15 and above 
who can, with understanding, read and write a short, simple statement on their 
everyday life. This dataset, sourced from UNESCO,  includes data from 1974 to 
2010.  

**Poverty Rates:**  
Gapminder defines extreme poverty as the percent of people who have less than 
$5.50 per day. This dataset, sourced from the World Bank, contains data from 
1966 to 2018. 

# 1. Literacy Rates by Decade
When looking at literacy rates grouped by decade, it is clear that countries are
becoming more literate, since the plots are moving further away from zero, and 
the density is becoming more clustered between a literacy rate between 90% and 
100%. This trend could be, in part, an artifact of the data, since each country 
is not represented in each decade, and the decade of 2000 has nearly twice the 
number of data points as any other decade. This trend that countries are 
becoming more literate could be investigated with more data and analysis further
to confirm its validity. 
```{r, echo=FALSE}
rates_violin(lit_gather)
```

# 2. Comparing Literacy Rates and Poverty Rates
Figure 2 compares a country's literacy rate with its poverty rate in a given 
year, and points are colored based on their region (note: the plot is 
interactive, so you can hover over a point to see more information). As we might
expect, European countries dominate the high literacy, low poverty region of the figure,while African countries are spread in the low literacy, high poverty 
region. In addition to Europe and Africa, each other region is grouped together, 
so, for example, most South American countries are other South American countries on the plot.   
One interesting pattern is that countries with a literacy rate below about 75% 
almost always have a poverty rate that is above 75%. In contrast, countries
with literacy rates above 75% have poverty rates that are spread from over 95% 
of individuals living in extreme poverty to a poverty rate of less than 1%. 


```{r, echo=FALSE}
rates_scatter(jn)
```