---
title: "Presentation"
author: "Rose Determan"
date: "10/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("wrangling_code.R")
source("visualization_function.R")
```
# Assignment description
1. describe your interesting observation seen within two subgroups of your data  
(think dplyr:filter() or maybe dplyr::group_by() and dplyr::summarise() )  
2. present your two comparative visualizations  
3. source the “wrangling_code.R” and the “visualization_function.R”  
4. will not display any visible code.   

I chose to examine **literacy rates and poverty rates** from the
[gapminder dataset](https://www.gapminder.org/data/). 
# Literacy Rates
Gapminder defines literacy rate as the percentage of people ages 15 and above 
who can, with understanding, read and write a short, simple statement on their 
everyday life. This dataset, sourced from UNESCO,  includes data from 1974 to 
2010.

# Poverty Rates 
Gapminder defines extreme poverty as the percent of people who have less than 
$5.50 per day. This dataset, sourced from the World Bank, contains data from 
1966 to 2018. 

# 1. Literacy Rates by Decade
```{r, echo=FALSE}
rates_violin(lit_dec)
```