---
title: "Global Literacy and Poverty Rates"
author: "Rose Determan"
date: "10/8/2021"
output:
  revealjs::revealjs_presentation:
    theme: simple
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("wrangling_code.R")
source("visualization_function.R")
```
<style type="text/css">
p { text-align: left; }
</style>

<style>
    .reveal section h1 {
    display: inline-block;
    font-size: 2em;
    line-height: 1.2em;
    vertical-align: top;
  }
</style>

<style>
    .reveal section p {
    font-size: 0.7em;
    line-height: 1.2em;
  }
</style>

<style>
    .reveal section li {
    font-size: 0.7em;
    line-height: 1.0em;
    vertical-align: top;
  }
</style>

# Dataset Introduction

**Literacy Rates:**  the percent of people ages 15 and above 
who can, with understanding, read and write a short, simple statement on their 
everyday life  
  
**Extreme Poverty Rates:** the percent of people who have less than 
$5.50 per day

# 1. Literacy Rates by Decade
- countries are becoming more literate, 
- perhaps an artifact of the data, since each country is not represented in each decade
```{r, echo=FALSE,fig.width=9,fig.height=4.5}
rates_violin(lit_gather)
```

# 2. Comparison Scatter Plot
<small>(note: the plot is interactive, so you can hover over a point to see more information)</small>

```{r, echo=FALSE,fig.width=9,fig.height=4.5}
rates_scatter(jn)
```

# 2. Comparison Plot Details
 - European countries dominate the high literacy, low poverty
 - African countries are spread in the low literacy, high poverty region.
 - Regions are grouped
 
Countries with a literacy rate below about 75%  almost always have a poverty rate that is above 75%. In contrast, countries with literacy rates above 75% have poverty rates that are spread from over 95% 
of individuals living in extreme poverty to a poverty rate of less than 1%.

# 3. Comparison Boxplots
- If a region has a high literacy rate, they the potential to have a low poverty rate. 
- Again, in regions where there is a high literacy rate, there can still be a high poverty rate.   
<small>Boxplots are sorted left to right from highest median literacy rate to lowest median literacy rate</small>
```{r, echo=FALSE,fig.width=9,fig.height=4.5, message =FALSE, warning = FALSE}
boxes(new)
```
